\sysfull

### 2018-08-05 paper/asplos19/intro.tex

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated, and may leave checkpointed volatile state inconsistent with non-volatile state. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, often leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based computer architecture support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are expensive to design, verify, and manufacture. New architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. \note{Talk about CP-based systems and talk why we focused task-based instead...}Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.

\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-v3.pdf}
    \caption{\note{small static task subplot and large static task subplot can be re-arranged to fit next to each other, so we end up with two layers plot} \note{Add distinction: top---legacy task-based system; bottom---Coala} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{figure}
%
%\textbf{Task Decomposition of Intermittent Programs.}

\textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks. A \emph{task}, a top level function, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
The programmer (or a compiler) explicitly expresses task-to-task control flow. Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead. At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at runtime. A programmer may thus create very large tasks in an attempt to reduce task transitions overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy buffer can hold which may lead to a task \emph{non-termination} problem. To eliminate this risk, existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress. These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.

>>>>

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated, and may leave checkpointed volatile state inconsistent with non-volatile state. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, often leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based computer architecture support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are expensive to design, verify, and manufacture. New architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. \note{Talk about CP-based systems and talk why we focused task-based instead...}Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-horiz.pdf}
    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}). On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
    \label{fig:coalesce}
\end{figure}
%
%\textbf{Task Decomposition of Intermittent Programs.} 
%
\textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks. A \emph{task}, a top level function, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
The programmer (or a compiler) explicitly expresses task-to-task control flow. Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead. At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}. The more task transitions a program requires, the more commit overhead is incurred by the system at runtime. A programmer may thus create very large tasks in an attempt to reduce task transitions overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy buffer can hold which may lead to a task \emph{non-termination} problem. To eliminate this risk, existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress. These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
